---
title: "Data preparation"
output:
  pdf_document: default
---

# Instructions

- You only need to submit the .Rmd of this file, not a PDF.

- You should __comment__ your code clearly to show what you've done to prepare the data.

- The purpose of this file is to use the data in the `data-raw` folder to create the data you will use in the report. The data you will use in the report should be saved in the `data` folder. It is good professional practice to make sure you're never directly modifying your raw data, but instead creating new datasets based on merges/manipulations that you need to reuse.

- Make sure you've taken a look at the hints for the web scraping and census API. 

- You may find the `write_rds()` function from the `readr` package helpful (it is loaded as part of the `tidyverse`).

- You do not need to keep the structure below.

# Set up

```{r, libraries, message=FALSE, echo=FALSE}
# install.packages("cancensus")
library(tidyverse)
library(janitor)
library(polite)
library(rvest)
library(haven)
library(cancensus)
library(sf)
library(geojsonsf)
library(lubridate)
library(gridExtra)
```


# Loading client data

## Reading From Raw

```{r, message=FALSE}
cust_dev_raw <- read_rds("data-raw/cust_dev.Rds")
cust_sleep_raw <- read_rds("data-raw/cust_sleep.Rds")
customer_raw <- read_rds("data-raw/customer.Rds")
device_raw <- read_rds("data-raw/device.Rds")
```

## Cleaning Data
```{r, message=FALSE}
# cleaning names with janitor
cust_dev_clean <- cust_dev_raw %>% 
  janitor::clean_names()

cust_sleep_clean <- cust_sleep_raw %>% 
  janitor::clean_names() 

customer_clean <- customer_raw %>% 
  janitor::clean_names() 

device_clean <- device_raw %>% 
  janitor::clean_names() 
```

```{r, message=FALSE}
# mutating na values numeric columns to 0
cust_dev_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

cust_sleep_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

customer_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

device_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)
```

```{r, message=FALSE}
# cleaning NA values - if any exists
cust_dev <- cust_dev_clean %>%
  filter(!is.na(cust_id) | !is.na(dev_id))
  
cust_sleep <- cust_sleep_clean %>% 
  filter(!is.na(cust_id) | !is.na(date) | !is.na(duration) | !is.na(flags))
  
customer <- customer_clean %>% 
  filter(!is.na(cust_id) | !is.na(dob) | !is.na(postcode) | !is.na(sex) |
           !is.na(pronouns) | !is.na(emoji_modifier))
  
device <- device_clean %>% 
  filter(!is.na(dev_id) | !is.na(device_name) | !is.na(line) | !is.na(released))
```


# Getting external data

## Web scraping industry data

```{r, message=FALSE}
url <- "https://fitnesstrackerinfohub.netlify.app/"

# Make sure this code is updated appropriately to provide 
# informative user_agent details
target <- bow(url,
              user_agent = "mrodin.karadeniz@utoronto.ca for STA303/1002 project",
              force = TRUE)

html <- scrape(target)

device_data_raw <- html %>% 
  html_elements("table") %>% 
  html_table() %>% 
  pluck(1) # added, in case you're getting a list format

# write scraped data to data-raw
write_rds(device_data_raw, "data-raw/device_data.rds")
```

## Cleaning scraped data
```{r, message=FALSE}
# cleaning names with janitor
device_data_clean <- device_data_raw %>% 
  janitor::clean_names()

# mutating na values numeric columns to 0
device_data_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

# cleaning NA id values - if any exists
device_data <- device_data_clean %>% filter(!is.na(device_name) |
                                            !is.na(line) |
                                            !is.na(recommended_retail_price) |
                                            !is.na(battery_life) |
                                            !is.na(water_resitance) |
                                            !is.na(heart_rate_sensor) |
                                            !is.na(pulse_oximiter) |
                                            !is.na(gps) |
                                            !is.na(sleep_tracking) |
                                            !is.na(smart_notifications) |
                                            !is.na(contactless_payments) |
                                            !is.na(released) |
                                            !is.na(brand))

```


# Census API

```{r, message=FALSE}
options(cancensus.api_key = "",
        cancensus.cache_path = "cache") # this sets a folder for your cache

# get all regions as at the 2016 Census (2020 not up yet)
regions <- list_census_regions(dataset = "CA16")

regions_filtered <-  regions %>% 
  filter(level == "CSD") %>% # Figure out what CSD means in Census data
  as_census_region_list()

# This can take a while
# We want to get household median income
census_data_csd <- get_census(dataset='CA16', regions = regions_filtered,
                          vectors=c("v_CA16_2397"), 
                          level='CSD', geo_format = "sf")

# Simplify to only needed variables
median_income_raw <- census_data_csd %>% 
  as_tibble() %>% 
  select(CSDuid = GeoUID, contains("median"), Population) %>% 
  mutate(CSDuid = parse_number(CSDuid)) %>% 
  rename(hhld_median_inc = 2)

# write census data to data-raw
write_rds(median_income_raw, "data-raw/median_income.rds")

```

## Cleaning census data
```{r, message=FALSE}
# cleaning names with janitor
median_income_clean <- median_income_raw %>% 
  janitor::clean_names()

# mutating na values numeric columns to 0
median_income_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

# cleaning NA id values - if any exists
median_income <- median_income_clean %>% 
  filter(!is.na(cs_duid) | !is.na(hhld_median_inc) | !is.na(population))

```

# Break glass in case of emergency
```{r, message=FALSE}
# read
postcode_conversion_raw <-
  read_rds("data-raw/break_glass_in_case_of_emergency.Rds")

# cleaning names with janitor
postcode_conversion_clean <- postcode_conversion_raw %>% 
  janitor::clean_names()

# cleaning NA id values - if any exists
postcode_conversion_clean %>% 
  filter(!is.na(pc) | !is.na(cs_duid))

# mutating na values numeric columns to 0
postcode_conversion <- postcode_conversion_clean %>% 
  mutate_if(is.numeric, replace_na, replace = 0)

```


# Data Mutations and Merges for Question 1
```{r, message=FALSE}
# Load the income of each costumer assume their income is the median income of 
# their neighborhood
customer_income <- customer
customer_income <- customer_income %>% rename(cs_duid = postcode)
customer_income$cs_duid <- 
  postcode_conversion$cs_duid[match(customer_income$cs_duid,
                                    postcode_conversion$pc)]

customer_income <- customer_income %>% add_column(income = "income")

customer_income$income <- median_income$hhld_median_inc[match(
  customer_income$cs_duid, median_income$cs_duid)]

# Calculate age from dob and drop dob
customer_info <- customer_income
customer_info$age = as.numeric(difftime(Sys.Date(), customer_info$dob, 
                                        units="weeks"))/52.25
customer_info <- select(customer_info, -dob)

# Rename emoji to represent skin color (whiteness) 1 being the whitest
# 5 being the darkest
customer_info <- customer_info %>% rename(skin_tone = emoji_modifier)
customer_info <- customer_info %>% 
  mutate(skin_tone = recode(
    skin_tone, "U+1F3FB" = "1", "U+1F3FC" = "2",
    "U+1F3FD" = "3", "U+1F3FE" = "4", "U+1F3FF" = "5"))

# Factor gender variable
customer_info$sex <- factor(customer_info$sex, 
                            levels = c("Female", "Male", "Intersex"))

# Add the device each costumer is using
customer_info$device_id <- cust_dev$dev_id[match(
  customer_info$cust_id, cust_dev$cust_id)]
customer_info$device <- device$device_name[match(customer_info$device_id, 
                                                 device$dev_id)]
customer_info$dev_line <- device$line[match(customer_info$device, 
                                                 device$device_name)]
customer_info <- select(customer_info, -device_id)

# Is the customer a newer one? 1 is yes, 0 is no
customer_info <- customer_info %>% 
  mutate(new = ifelse(dev_line == "Active" , "1", 
                      ifelse(dev_line == "Advance", "1", "0")))
customer_info$new <- as.numeric(customer_info$new)
customer_info <- na.omit(customer_info)

# New customer data (customer who uses advance and active)
new_cust <- customer_info[customer_info$new == 1,]
new_cust <- na.omit(new_cust)

```


# Data Mutations and Merges for Question 2
```{r, message=FALSE}
# We will now merge customer data and sleep data
cust_sleep_data <- right_join(customer, cust_sleep, by = "cust_id")

# Remove all observations that have NA values for emoji_modifier & Sex
cust_sleep_data <- cust_sleep_data %>% filter(!is.na(emoji_modifier)) %>%
  filter(!is.na(sex))

# Add rate of flags to data
cust_sleep_data <- cust_sleep_data %>%
  mutate(flag_rate = flags/ceiling(duration/60))

# Add a variable that generalizes the emoji skin to according to skin tone
cust_sleep_data <- cust_sleep_data %>%
  mutate(skin_tone = ifelse(emoji_modifier == "U+1F3FF", "dark", 
                            ifelse(emoji_modifier == "U+1F3FE", "medium dark",
                                   ifelse(emoji_modifier == "U+1F3FD", "medium",
                                          ifelse(emoji_modifier == "U+1F3FC",
                                                 "medium light", "light")))))

# Get customer age using their DOB
cust_sleep_data <- cust_sleep_data %>%
  mutate(age = as.numeric(2022 - year(dob)))

# Create age groups according to Statistics Canada Website
cust_sleep_data <- cust_sleep_data %>%
  mutate(age_group = ifelse(age >= 15 && age <= 24, "Youth",
                            ifelse(age >= 25 && age <= 64, "Adulthood",
                                   ifelse(age >= 65, "Senior", "Empty"))))

# Factor gender variable
cust_sleep_data$sex <- factor(cust_sleep_data$sex, 
                            levels = c("Female", "Male", "Intersex"))
```


# Writing to data
```{r, message=FALSE}
# Customer Data
write_rds(cust_dev, "data/cust_dev.rds")
write_rds(cust_sleep, "data/cust_sleep.rds")
write_rds(customer, "data/customer.rds")
write_rds(device, "data/device.rds")
write_rds(customer_info, "data/customer_info.rds")
write_rds(new_cust, "data/new_cust.rds")
write_rds(cust_sleep_data, "data/cust_sleep_data.rds")

# Web Scraping
write_rds(device_data, "data/device_data.rds")

# Census API
write_rds(median_income, "data/median_income.rds")

# Break glass in case of emergency
write_rds(postcode_conversion, "data/postcode_conversion_data.rds")

```


